---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

# Classification

## Lab: Logistic Regression, LDA, QDA, and KNN

### 4.6.2 Logistic Regression

Let's return to the `Smarket` data from `ISLR`. 

```{r}
library(ISLR)
library(dplyr)
names(Smarket)
dim(Smarket)
summary(Smarket)
```

In this lab, we will fit a logistic regression model in order to predict `Direction` using `Lag1` through `Lag5` and `Volume`. The `glm()` function fits **generalized linear models**, a class of models that includes logistic regression. 

The syntax of the `glm()` function is similar to that of `lm()`, except that we must pass in the argument `family=binomial` in order to tell `R` to run a logistic regression rather than some other type of generalized linear model.

```{r}
glm_fit = glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume, data=Smarket, family=binomial)
summary(glm_fit)
```

The smallest p-value here is associated with `Lag1`. The negative coefficient
for this predictor suggests that if the market had a positive return yesterday,
then it is less likely to go up today. However, at a value of 0.145, the p-value
is still relatively large, and so there is no clear evidence of a real association
between `Lag1` and `Direction`.

If you want, you can use helper methods like `coefficients()` and `residuals()` in order to access specific parts of this fitted model:

```{r}
coefficients(glm_fit)
residuals(glm_fit)
```

The `predict()` function can be used to predict the probability that the
market will go up, given values of the predictors. The `type="response"`
option tells `R` to output probabilities of the form `P(Y = 1|X)`, as opposed
to other information such as the `logit`. If no data set is supplied to the
`predict()` function, then the probabilities are computed for the training
data that was used to fit the logistic regression model. 

```{r}
glm_probs = data.frame(probs = predict(glm_fit, type="response"))
head(glm_probs)
```

Here we have printed only the first few probabilities. We know that these values correspond to the probability of the market going up, rather than down, because the
`contrasts()` function indicates that `R` has created a dummy variable with
a 1 for `Up`.

```{r}
contrasts(Smarket$Direction)
```

In order to make a prediction as to whether the market will go up or
down on a particular day, we must convert these predicted probabilities
into class labels, `Up` or `Down`. The following two commands create a variable with 
class predictions based on whether the predicted probability of a market
increase is greater than or less than 0.5.

```{r}
glm_pred = glm_probs %>%
  mutate(pred = ifelse(probs>.5, "Up", "Down"))
```

The second line transforms to `Up` all of the elements for which the predicted probability of a
market increase exceeds 0.5 and to `Down` all of those less than or equal to 0.5.
Given these predictions, we can `count()` how many observations were correctly or incorrectly classified.

```{r}
glm_pred = cbind(Smarket, glm_pred)

glm_pred %>% 
  count(class, Direction)

glm_pred %>%
  summarize(score = mean(pred == Direction))
```

Our model correctly predicted that the market would go up on 507 days and that
it would go down on 145 days, for a total of 507 + 145 = 652 correct
predictions. The `summarize()` function can be used to compute the fraction of
days for which the prediction was correct. In this case, logistic regression
correctly predicted the movement of the market 52.2% of the time.

At first glance, it appears that the logistic regression model is working
a little better than random guessing. But remember, this result is misleading
because we trained and tested the model on the same set of 1,250 observations.
In other words, 100− 52.2 = 47.8% is the **training error rate**. As we
have seen previously, the training error rate is often overly optimistic — it
tends to underestimate the _test_ error rate. 

In order to better assess the accuracy
of the logistic regression model in this setting, we can fit the model
using part of the data, and then examine how well it predicts the held out
data. This will yield a more realistic error rate, in the sense that in practice
we will be interested in our model’s performance not on the data that
we used to fit the model, but rather on days in the future for which the
market’s movements are unknown.

Like we did with KNN, we will first create a vector corresponding
to the observations from 2001 through 2004. We will then use this vector
to create a held out data set of observations from 2005.

```{r}
train = Smarket %>%
  filter(Year < 2005)

test = Smarket %>%
  filter(Year >= 2005)
```

We now fit a logistic regression model using only the subset of the observations
that correspond to dates before 2005, using the subset argument.
We then obtain predicted probabilities of the stock market going up for
each of the days in our test set—that is, for the days in 2005.

```{r}
glm_fit = glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume, data=train, family=binomial)
glm_probs = data.frame(probs = predict(glm_fit, newdata=test, type="response"))
```

Notice that we have trained and tested our model on two completely separate
data sets: training was performed using only the dates before 2005,
and testing was performed using only the dates in 2005. Finally, we compute
the predictions for 2005 and compare them to the actual movements
of the market over that time period.

```{r}
glm_pred = glm_probs %>%
  mutate(pred = ifelse(probs>.5, "Up", "Down"))

glm_pred = cbind(test, glm_pred)

glm_pred %>% 
  count(class, Direction)

glm_pred %>%
  summarize(score = mean(pred == Direction),
            recip = mean(pred != Direction))
```

The `!=` notation means **not equal to**, and so the last command computes
the test set error rate. The results are rather disappointing: the test error
rate is 52%, which is worse than random guessing! Of course this result
is not all that surprising, given that one would not generally expect to be
able to use previous days’ returns to predict future market performance.
(After all, if it were possible to do so, then the authors of this book [along with your professor] would probably
be out striking it rich rather than teaching statistics.)

We recall that the logistic regression model had very underwhelming pvalues
associated with all of the predictors, and that the smallest p-value,
though not very small, corresponded to `Lag1`. Perhaps by removing the
variables that appear not to be helpful in predicting `Direction`, we can
obtain a more effective model. After all, using predictors that have no
relationship with the response tends to cause a deterioration in the test
error rate (since such predictors cause an increase in variance without a
corresponding decrease in bias), and so removing such predictors may in
turn yield an improvement. 

In the space below, refit a logistic regression using just `Lag1` and `Lag2`, which seemed to have the highest predictive power in the original logistic regression model.

```{r}
glm_fit = # Write your code to fit the new model here

# This will test your new model
glm_probs = data.frame(probs = predict(glm_fit, newdata=test, type="response"))

glm_pred = glm_probs %>%
  mutate(pred = ifelse(probs>.5, "Up", "Down"))

glm_pred = cbind(test, glm_pred)

glm_pred %>% 
  count(class, Direction)

glm_pred %>%
  summarize(score = mean(pred == Direction))
```

Now the results appear to be more promising: 56% of the daily movements
have been correctly predicted. The confusion matrix suggests that on days
when logistic regression predicts that the market will decline, it is only
correct 50% of the time. However, on days when it predicts an increase in
the market, it has a 58% accuracy rate.

Finally, suppose that we want to predict the returns associated with **particular
values** of `Lag1` and `Lag2`. In particular, we want to predict Direction on a
day when `Lag1` and `Lag2` equal 1.2 and 1.1, respectively, and on a day when
they equal 1.5 and −0.8. We do this using the `predict()` function.

```{r}
predict(glm_fit, newdata=data.frame(Lag1=c(1.2,1.5), Lag2=c(1.1,-0.8)), type="response")
```

To get credit for this lab, play around with a few other values for `Lag1` and `Lag2`, and then post to Piazza about what you found. If you're feeling adventurous, try fitting models with other subsets of variables to see if you can find a better one!

### 4.6.3 Linear Discriminant Analysis

Now we will perform LDA on the `Smarket` data from the `ISLR` package. In `R`, we can fit a LDA model using the `lda()` function, which is part of the `MASS` library. **Note:** `dplyr` and `MASS` have a name clash around the word `select()`, so we need to do a little magic to make them play nicely.

```{r}
library(MASS)
library(dplyr)
library(ISLR)
select <- dplyr::select
```

The syntax for the `lda()` function is identical to that of `lm()`, and to that of
`glm()` except for the absence of the `family` option. As we did with logistic regression and KNN, we'll fit the model using only the observations before 2005, and then test the model on the data from 2005.

```{r}
train = Smarket %>%
  filter(Year < 2005)

test = Smarket %>%
  filter(Year >= 2005)

model_LDA=lda(Direction~Lag1+Lag2, data=train)
print(model_LDA)
```

The LDA output indicates prior probabilities of ${\hat{\pi}}_1 = 0.492$ and ${\hat{\pi}}_2 = 0.508$; in other words,
49.2% of the training observations correspond to days during which the
market went down.

The function also provides the group means; these are the average
of each predictor within each class, and are used by LDA as estimates
of $\mu_k$. These suggest that there is a tendency for the previous 2 days’
returns to be negative on days when the market increases, and a tendency
for the previous days’ returns to be positive on days when the market
declines. 

The coefficients of linear discriminants output provides the linear
combination of `Lag1` and `Lag2` that are used to form the LDA decision rule.

If $−0.642\times{\tt Lag1}−0.514\times{\tt Lag2}$ is large, then the LDA classifier will
predict a market increase, and if it is small, then the LDA classifier will
predict a market decline. 

We can use the `plot()` function to produce plots of the linear
discriminants, obtained by computing $−0.642\times{\tt Lag1}−0.514\times{\tt Lag2}$ for
each of the training observations.

```{r}
plot(model_LDA)
```

The `predict()` function returns a list with three elements. The first element,
`class`, contains LDA’s predictions about the movement of the market.
The second element, `posterior`, is a matrix whose $k^{th}$ column contains the
posterior probability that the corresponding observation belongs to the $k^{th}$
class. Finally, `x` contains the linear discriminants,
described earlier.

```{r}
predictions_LDA = data.frame(predict(model_LDA, test))
names(predictions_LDA)
```

Let's check out the confusion matrix to see how this model is doing. We'll want to compare the **predicted class** (which we can find in the `class` column of the `predictions_LDA` data frame) to the **true class**.

```{r}

predictions_LDA = cbind(test, predictions_LDA)

predictions_LDA %>%
  count(class, Direction)

predictions_LDA %>%
  summarize(score = mean(class == Direction))
```

The LDA predictions are identical to the ones from our logistic model:

```{r}
# Logistic model, for comparison
model_logistic = glm(Direction~Lag1+Lag2, data=train ,family=binomial)

logistic_probs = data.frame(probs = predict(model_logistic, test, type="response"))

predictions_logistic = logistic_probs %>%
  mutate(class = ifelse(probs>.5, "Up", "Down"))

predictions_logistic = cbind(test, predictions_logistic)

predictions_logistic %>%
  count(class, Direction)

#glm_pred %>%
predictions_logistic %>%
  summarize(score = mean(class == Direction))
```


### 4.6.4 Quadratic Discriminant Analysis
We will now fit a QDA model to the `Smarket` data. QDA is implemented
in `R` using the `qda()` function, which is also part of the MASS library. The
syntax is identical to that of `lda()`.

```{r}
model_QDA = qda(Direction~Lag1+Lag2, data=train)
model_QDA
```

The output contains the group means. But it does not contain the coefficients
of the linear discriminants, because the QDA classifier involves a
_quadratic_, rather than a linear, function of the predictors. The `predict()`
function works in exactly the same fashion as for LDA.

```{r}
predictions_QDA = data.frame(predict(model_QDA, test))

predictions_QDA = cbind(test, predictions_QDA)

predictions_QDA %>%
  count(class, Direction)

predictions_QDA %>%
  summarize(score = mean(class == Direction))
```

Interestingly, the QDA predictions are accurate almost 60% of the time,
even though the 2005 data was not used to fit the model. This level of accuracy
is quite impressive for stock market data, which is known to be quite
hard to model accurately. 

This suggests that the quadratic form assumed
by QDA may capture the true relationship more accurately than the linear
forms assumed by LDA and logistic regression. However, we recommend
evaluating this method’s performance on a larger test set before betting
that this approach will consistently beat the market!

### An Application to Carseats Data

Let's see how the `LDA/QDA` approach performs on the `Carseats` data set, which is
part of the `ISLR` library. 

Recall: this is a simulated data set containing sales of child car seats at 400 different stores.

```{r}
summary(Carseats)
```

See if you can build a model that predicts `ShelveLoc`, the shelf location (`Bad`, `Good`, or `Medium`) of the product at each store. Don't forget to hold out some of the data for testing!

```{r}
# Your code here
```

To get credit for this lab, please post your answers to the following questions:

- What was your approach to building the model?
- How did your model perform?
- Was anything easier or more challenging than you anticipated?

to Piazza: https://piazza.com/class/isuhyl9uxya5vm?cid=14
### 4.6.5: K-Nearest Neighbors

In this lab, we will perform KNN on the `Smarket` dataset from `ISLR`. This data set consists of percentage returns for the S&P 500 stock index over 1,250 days, from the
beginning of 2001 until the end of 2005. For each date, we have recorded
the percentage returns for each of the five previous trading days, `Lag1` through `Lag5`. We have also recorded `Volume` (the number of shares traded on the previous day, in billions), `Today` (the percentage return on the date in question) and `Direction` (whether the market was `Up` or `Down` on this date). We can use the `slice()` function to look at the first few rows:

```{r, message=FALSE}
library(ISLR)
library(tidyverse)
library(mosaic)
data(Smarket)
Smarket %>% slice(1:10)
```

Today we're going to try to predict ` Direction` using percentage returns from the previous two days (`Lag1` and `Lag2`). We'll build our model using the `knn()` function, which is part of the
`class` library:

```{r}
library(class)
```

This function works rather differently from the other model-fitting
functions that we have encountered thus far. Rather than a two-step
approach in which we first fit the model and then we use the model to make
predictions, `knn()` forms predictions using a single command. The function
requires four inputs.
   1. Training data (just the predictors). We'll call this `train_Market`.
   2. Testing data (just the predictors). We'll call this `test_Market`.
   3. Training data (our outcome variable, which is class labels in this case). We'll call this `train_Direction`.
   4. A value for $K$, the number of nearest neighbors to be used by the classifier.

We'll first create two subsets of our data-- one containing the observations from 2001 through 2004, which we'll use to train the model and one with observations from 2005 on, for testing. To do this, we'll use the `dplyr` `filter()` command. 

```{r}
train = Smarket %>%
  filter(Year < 2005)
test = Smarket %>%
  filter(Year >= 2005)
```


For the model, we need small datasets for `train_Market` and `test_Market` that only contain the predictors we want to use. Since we're interested in `Lag1` and `Lag2`, we need to pull those out. We can do this using `select()`. 

```{r}
train_Market = train %>%
  select(Lag1, Lag2)
test_Market = test %>%
  select(Lag1, Lag2)
```

Because of the way the `knn()` function is written, we need to provide the class labels as a vector, which requires an additional step to the pipeline. By adding `.$Direction` to the end of our `dplyr` chain, we're indicating that we want to turn the result of our `select()` (which is a dataframe, as `dplyr` makes all its outputs) into a vector. 

```{r}
train_Direction = train %>%
  select(Direction) %>%
  .$Direction 
```

Now the `knn()` function can be used to predict the market’s movement for
the dates in 2005. We set a **random seed** before we apply `knn()` because
if several observations are tied as nearest neighbors, then `R` will randomly
break the tie. Therefore, a seed must be set in order to ensure reproducibility
of results. You can put in your favorite number here, or leave it as 1. 

```{r}
set.seed(1)
knn_pred = knn(train_Market, test_Market, train_Direction, k=1)
```

The `table()` function can be used to produce a **confusion matrix** in order to determine how many observations were correctly or incorrectly classified.

```{r}
test_Direction = test %>%
  select(Direction) %>%
  .$Direction
table(knn_pred, test_Direction)
mean(knn_pred==test_Direction)
```

The results using $K = 1$ are not very good, since only 50% of the observations
are correctly predicted. Of course, it may be that $K = 1$ results in an
overly flexible fit to the data. Below, we repeat the analysis using $K = 3$.

```{r}
knn_pred3 = knn(train_Market, test_Market, train_Direction, k=3)
table(knn_pred3, test_Direction)
mean(knn_pred3==test_Direction)
```

The results have improved slightly. Let's try a few other $K$ values to see if we get any further improvement. To do this, we're going to use a few more of the tidyverse packages, `tidyr` and `purrr`. These are automically loaded by the convenience package `tidyverse`, so we don't need to load them explicitly. But, it's still good to know where functions come from so you can search for help effectively!

```{r}
set.seed(1)

predictions = data_frame(k = 1:5) %>%
  unnest(prediction = map(k, ~ knn(train_Market, test_Market, train_Direction, k = .))) %>%
  mutate(oracle = rep_along(prediction, test_Direction))

predictions %>%
  group_by(k) %>%
  summarize(accuracy = mean(prediction == oracle))
```

It looks like for classifying this dataset, KNN might not be the right approach.

### 4.6.6: An Application to Caravan Insurance Data
Let's see how the `KNN` approach performs on the `Caravan` data set, which is
part of the `ISLR` package. This data set includes 85 predictors that measure demographic characteristics for 5,822 individuals. The response variable is `Purchase`, which indicates whether or not a given individual purchases a
caravan insurance policy. In this data set, only 6% of people purchased
caravan insurance.

```{r}
tally(~Purchase, data=Caravan, format = "percent")
```

Because the `KNN` classifier predicts the class of a given test observation by
identifying the observations that are nearest to it, the scale of the variables
matters. Any variables that are on a large scale will have a much larger
effect on the distance between the observations, and hence on the `KNN`
classifier, than variables that are on a small scale. 

For instance, imagine a
data set that contains two variables, salary and age (measured in dollars
and years, respectively). As far as `KNN` is concerned, a difference of \$1,000
in salary is enormous compared to a difference of 50 years in age. Consequently,
salary will drive the `KNN` classification results, and age will have
almost no effect. 

This is contrary to our intuition that a salary difference
of \$1,000 is quite small compared to an age difference of 50 years. Furthermore,
the importance of scale to the `KNN` classifier leads to another issue:
if we measured salary in Japanese yen, or if we measured age in minutes,
then we’d get quite different classification results from what we get if these
two variables are measured in dollars and years.

A good way to handle this problem is to **standardize** the data so that all
variables are given a mean of zero and a standard deviation of one. Then
all variables will be on a comparable scale. The `scale()` function does just
this. In standardizing the data, we exclude the qualitative `Purchase` variable.

```{r}
standardized_Caravan = Caravan %>%
  select(-Purchase) %>%
  scale() %>%
  data.frame() # This is an artifact of using the base scale() function. There must be a tidyier way!

var(~MOSTYPE, data=Caravan)
var(~MAANTHUI, data=Caravan)

var(~MOSTYPE, data=standardized_Caravan)
var(~MAANTHUI, data=standardized_Caravan)
```

Now every column of `standardized_Caravan` has a standard deviation of one and
a mean of zero.

We'll now split the observations into a test set, containing the first 1,000
observations, and a training set, containing the remaining observations.

```{r}
test_Caravan = standardized_Caravan %>%
  slice(1:1000)
train_Caravan = standardized_Caravan %>%
  slice(1001:5822)

Purchase = Caravan %>%
  select(Purchase)

test_Purchase = Purchase %>%
  slice(1:1000) %>%
  .$Purchase

train_Purchase = Purchase %>%
  slice(1001:5822) %>%
  .$Purchase
```

Let's fit a `KNN` model on the training data using $K = 1$, and evaluate its
performance on the test data.

```{r}
set.seed(1)
knn_pred = knn(train_Caravan, test_Caravan, train_Purchase, k=1)
mean(test_Purchase != knn_pred) # KNN error rate
mean(test_Purchase != "No")     # Percent of people who purchase insurance
```

The KNN error rate on the 1,000 test observations is just under 12%. At first glance, this may appear to be fairly good. However, since only 6% of customers purchased insurance, we could get the error rate down to 6% by always predicting `No` regardless of the values of the predictors!

Suppose that there is some non-trivial cost to trying to sell insurance
to a given individual. For instance, perhaps a salesperson must visit each
potential customer. If the company tries to sell insurance to a random
selection of customers, then the success rate will be only 6%, which may
be far too low given the costs involved. 

Instead, the company would like
to try to sell insurance only to customers who are likely to buy it. So the
overall error rate is not of interest. Instead, the fraction of individuals that
are correctly predicted to buy insurance is of interest.

It turns out that `KNN` with $K = 1$ does far better than random guessing
among the customers that are predicted to buy insurance:

```{r}
table(knn_pred, test_Purchase)
```

Among 77 such customers, 9, or 11.7%, actually do purchase insurance. This is double the
rate that one would obtain from random guessing. Let's see if increasing $K$ helps! Try out a few different $K$ values below. Feeling adventurous? Write some code that figures out the best value for $K$.

```{r}
set.seed(1)
# Your code here
```

It appears that `KNN` is finding some real patterns in a difficult data set! To get credit for this lab, post a response to the Piazza prompt available at: https://piazza.com/class/igwiv4w3ctb6rg?cid=10